Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-01-09 manuscript.Rnw
Refine author list
Reworking of main text

\author[1]{Remko A. Duursma}
\author[2]{Daniel S. Falster}
\author[3]{the BAAD consortium}

\affil[1]{Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW 2751, Australia}
\affil[2]{Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}
\affil[3]{See Supporting Information}

@@

\textbf{The partitioning of biomass into leaves and stems is one of the most uncertain  and influential components of  global vegetation models (GVMs), affecting the residence time of carbon in ecosystems \citep{friend2014}, predictions of global carbon storage \citep{ise2010}, and the CO\textsubscript{2} response of net primary production \citep{dekauwe2014}. Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new global database on biomass partitioning from over 175 studies, we here show a strong effect of PFT on partitioning between stems and leaves. We show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum \citep{wrightetal2004}. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current GVMs.
}

@@

Globally, forests store ca. 360 Pg C in live biomass \citep{panetal2011}, or about 40 years of current athropogenic emissions \citep{friedlingsteinetal2014}. The partitioning of this biomass among plant components is a major determinant of vegetation productivity and the rate of carbon cycling \cite{ise2010}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008},  thus representing a significant long-term carbon storage component. By contrast, leaf biomass turns over quickly, and is a major input to the soil carbon cycle where the majority of carbon is released back to the atmosphere. Recently, the residence time of carbon in vegetation has been identified as a major source of uncertainty in model projections of effects of climate change on the global carbon sink \citep{friend2014}. There is little consensus in these global vegetation models (GVMs) on how biomass partitioning should be modelled \citep{franklin2012,dekauwe2014}, but there is no shortage of underlying theory \citep{cannelldewar1994,franklin2012}. We stress that biomass partitioning, the actual distribution of biomass among components, is distinct from biomass allocation, the flux of carbon into plant components. Biomass partitioning is estimated from allocation by including the rates of turnover of the various plant components (see Methods S1).

It is difficult to decide which partitioning models are superior, since GVMs are usually only tested against productivity or carbon storage data, and both are the outcome of multiple processes. The vast majority of GVMs simplify parameterization by considering a few plant functional types (PFTs) \citep{harrisonetal2010}. It is also common to assume that parameters related to biomass partitioning varies between PFTs, however, parameterization of these differences have been based on little empirical evidence.  Based on previous compilations of forest biomass estimates, it is known that gymnosperms have more leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF) than angiosperms \citep{oneilldeangelis1981,enquistniklas2002,reich2014}. However, it is not known whether this is due to the angiosperm/evergreen divergence, or to inherent differences in leaf life span. If the latter is true, we would expect evergreen angiosperms to have more leaf biomass than deciduous angiosperms.

We argue that the lack of a large global database on biomass partitioning at the individual plant level has hampered progress in this field. Here, we use a unique database to lay the empirical base against which models can be benchmarked. Our new database (BAAD, Falster) includes over 20,000 observations contributed by 175 studies of biomass and size metrics for individual plants, covering the entire size range including saplings and the largest trees in the world, three major woody PFTs, three biomes (tropical, temperate, and boreal forests), and nine vegetation types (Fig. S1). Another unique aspect is that the database contains leaf mass as well as leaf area estimates, whereas previous large compilations only considered leaf mass \citep{oneilldeangelis1981,enquistniklas2002,reich2014}. It is unclear whether PFTs with higher leaf biomass will also have higher total leaf area, because PFTs differ in the ratio of leaf mass to area (LMA), a measure of the leaf construction cost \citep{poorter2009}. Leaf area itself is relevant because light interception acts on leaf area, which drives photosynthesis. A few classic studies have assumed that gymnosperms indeed carry much more leaf area than angiosperms \citep{chabothicks1982, bond1989}. Finally it is also unknown whether potential differences between PFTs in biomass partitioning are a result of inherent differences in morphology, or whether climate directly affects partitioning, and the PFTs simply occupy different climate spaces.

We hypothesize that gymnosperms have higher LMF than angiosperms, and that evergreen angiosperms have higher LMF than deciduous angiosperms. We further test whether differences between PFTs are related to LMA. We use several variables from BAAD (Fig. 1a) to test this main hypothesis. We also test whether climate has a direct influence on biomass partitioning. This is possible because the database has very good global coverage of the climate space (Fig. 1b).

We found that LMF differed strongly between the three PFTs tested, with evergreen gymnosperms > evergreen angiosperms > deciduous angiosperms (Fig. 2a). Because of allometric relationships between the studied variables \citep{niklas1994}, there was a very strong effect of plant height on LMF (Fig 2a and S2). We corrected for plant height by estimating LMF at a common plant height, using a generalized additive model, a semi-parametric method that makes no assumption on the underlying functional relationships \citep{wood2006}. We then found that LMF at a common plant height was proportional to the leaf construction cost, as measured by leaf mass per area (Fig. 2b). The consequence of this is that the leaf area ratio (the amount of leaf area supported per unit above-ground biomass), was invariant between PFTs, because $LAR = LMF / LMA$ (Fig 2c and S3). We thus supported the hypothesis that PFTs differ in LMF, and for the first time demonstrated that they converge in the amount of leaf area supported.

Next, we decomposed LMF into two components, $LMF = M_F/M_T = M_F/A_S * A_S/M_T$, where $M_F/A_S$ is the ratio of leaf biomass to stem basal area (see Methods S2), and $A_S/M_T$ the ratio of stem basal area to total above-ground biomass. We found that the first term showed identical PFT patterns as LMF (Fig 2d and e), $M_F/A_S$ was highly variable between PFT and proportional to LMA, so that $A_F/A_S$ is invariant between PFTs. This decomposition is also convenient because the first term is nearly independent of plant height (Tables 1 and 2), allowing a more straightforward analysis of climate effects (not having to correct for height every time). The second term ($A_S/MT$) was found to be nearly unaffected by PFT, but strongly height-dependent (Fig. S4). It thus absorbs variation related to height and above-ground plant geometry, and is consistent with previous findings that this ratio is fairly conserved between plant species \citep{chave2005}.

We analyzed climate effects in two ways: with quantitative variables (mean annual precipitation; MAP, mean annual temperature; MAT), and by biome (tropical / temperate/ boreal), which is a simple classification taking into account both MAT and MAP (Fig. S1). Regardless of how we analyzed the data, we found only weak climate effects. We did not find differences between biomes in $M_F/A_S$ (Fig. 3), but some small inconsistent differences in $A_F/A_S$ (corresponding to slightly lower LMA in tropical vs. temperate evergreen angiosperms). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table 2). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) on $M_F/A_S$ and $A_F/A_S$ (Fig. 4). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table 1). The only exception was LMA, where the $R^2$ increased from 0.47 to 0.61 when climate variables were added, consistent with previous reports that LMA shows higher plasticity than biomass partitioning \citep{poorter2012}.

We found that PFTs differed in LMF, and this was proportional to leaf mass per area (LMA), a leading trait within the leaf economics spectrum. In addition, LMA alone explained 28\% of the variation in $M_F/A_S$ across species, without accounting for PFT (Fig. S6) (and consistent with averages by PFT, $A_F/A_S$ was very weakly related to LMA, with an $R^2$ of 0.02). It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. One explanation for this proportionality may be the correlation between LMA and leaf lifespan \citep{wrightetal2004}. If annual leaf production is constant, longer leaf lifespan will simply result in the accumulation of more leaf cohorts, and thus biomass \citep{chabothicks1982,enquistniklas2002,poorter2012}. However, leaf lifespan (LL) is correlated to but not proportional to LMA ($LL \propto LMA^{0.58}$, from Ref. \citep{wrightetal2004}). Another interpretation is that plants may target a certain leaf area given a certain biomass investment in supporting tissues, because light interception and photosynthesis correlate with the exchange area of the plant.

We found very only weak and inconsistent effects of climate on biomass partitioning, in contrast to a recent study where LMF was found to be correlated with MAT \citep{reich2014}. One explanation for these differences is that they studied whole-stand biomass, not individual plant biomass. This begs the question whether climate effects on stand-level biomass partitioning are on stand density (number of individuals per unit area) or individual plants. Our finding that climate does not appreciably affect biomass partitioning is good news for models that parameterize the globe by PFT alone. However, it is inconsistent with studies demonstrating small-scale effects of climate on leaf area to sapwood area \citep{callaway1994,poyatosetal2007}, an important component of $A_F/A_S$. We do stress that there is still much variation between species and studies unaccounted for (Fig. S5), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning. Nonetheless, we can firmly conclude that effects of plant functional type were much stronger than direct climate effects (Tables 1 and 2).

A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. We argue that these models should be constrained by patterns with plant size and PFT presented in our study. As a first approximation, GVMS should aim to keep leaf area to stem cross-sectional area constant, and vary biomass allocation accordingly. The leaf mass per area, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem biomass. An example of such a model is LPJ \citep{sitchetal2003}, where the ratio of leaf area to sapwood area is a constant parameter, but LMA differs between PFTs. Models that use constant allocation fractions may also be able to mimic the patterns between PFTs observed here, but only if leaf life span differs in such a way to maintain a proportionality between LMF and LMA. In any case, it is relatively straightforward to compare model simulations to broad patterns in LMF with PFT and plant height as presented in this paper. The data we use here is freely available (Falster), which should aid model testing.

@@

We used the recently compiled Biomass And Allometry Database (BAAD) (Falster), including a total of \Sexpr{nrow(baad_all$data)} observations, but here we use only data that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_W$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (ca. 1.3m) ($A_{Sbh}$).  For each analysis, we used different subsets of the data. Sample sizes by PFT are summarized in Tables~\ref{table:count1}-\ref{table:count4} We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

$$LMF = M_F / M_T = (M_F / A_F) (A_F / M_T) = LMA * LAR$$

@@

$$LMF = M_F / M_T = (M_F / A_S) (M_T / A_S)^{-1}$$

@@

R.A.D and D.S.F. conceived the ideas and methods, and co-wrote the paper. R.A.D. analysed data. Co-authors in the BAAD consortium shared data.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{Figure3.pdf}
    \caption{\textbf{No biome differences in leaf mass per unit stem basal stem area for $M_F/A_S$ and weak, inconsistent effects of biome on leaf area per unit basal stem area ($A_F/A_S$)}. No leaf area estimates were available for boreal deciduous angiosperms. DA = Deciduous Angiosperm, EA = Evergreen Angiosperm, EG = Evergreen Gymnosperm. Error bars are 95\% confidence intervals for the mean, estimated from a linear mixed-effects model using species within study as the random effect. Letters denote significant differences (at $\alpha = 0.05$)}
    \label{fig:figure3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure4.pdf}
    \caption{\textbf{Modest effects of climate on leaf mass and area per unit stem basal area, compared to the very large effect of plant functional type on mass-based variables ($M_F/A_S$)}. Solid lines are generalized additive model (GAM) fits (with a base dimension of 3, larger values resulted in biologically implausible relationships). Grey areas are approximate 95\% confidence intervals, calculated by the GAM. To account for within-species and contributed dataset correlations, data were first averaged by species/dataset combination.}
  \label{fig:figure4}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS6.pdf}
    \caption{\textbf{Plant leaf mass as a function of basal stem area (left panel), and leaf area as a function of basal stem area (right panel).} Individual plant data were averaged by species/study (the same species occurring in multiple datasets thus appears multiple times). Lines are linear regressions with 95\% confidence intervals. Both regressions were significant ($P \textless 0.01$), but $R^2$ was 0.28 in the left panel, and only 0.02 in the right panel.}
  \label{fig:figureS6}
\end{figure}

@@

<<results='asis'>>=
print(xtable(dataauthors,
             caption="\\textbf{Co-authors in the BAAD consortium.}"
             ), include.rownames=FALSE, tabular.environment='longtable', floating=FALSE, size="small",
      caption.placement="top")
@

@@

$$
\frac{dM_i}{dt} = a_i NPP - \tau_i M_i
$$

>>>>

\author[1]{Remko A. Duursma}
\author[2]{Daniel S. Falster}

\affil[1]{Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW 2751, Australia}
\affil[2]{Biological Sciences, Macquarie University, North Ryde, NSW 2109, Australia}

@@

\textbf{The partitioning of biomass into leaves and stems is one of the most uncertain  and influential components of  global vegetation models (GVMs), affecting the residence time of carbon in ecosystems \citep{friend2014}, predictions of global carbon storage \citep{ise2010}, and the CO\textsubscript{2} response of net primary production \citep{dekauwe2014}. Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new global database on biomass partitioning from over 175 studies, we here show a strong effect of PFT on partitioning between stems and leaves. We show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum \citep{wrightetal2004}. In addition, we show how a second trait, leaf area per stem basal area, contributes additional variation in biomass partitioning among species; however this trait does not differ systematically among major woody PFTs. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current GVMs.
}

@@

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation\cite{ise2010}. Globally, forests store approximately 360Pg of carbon\citep{panetal2011}, equivalent to about 40years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Biomass stored in woody stems is unproductive but has a long residence time \citep{luyssaert2008}, whereas leaf biomass is enhances photosynthesis but turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \cite{ryanlaw2005}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014}.

Perhaps and the biggest challenge for GVMs is to capture the combined responses of the more than 250,000 plant species comprising world vegetation. While most plants have the same basic resource requirements and physiological function, large differences exist among species in the amount of energy invested in different tissues (leaves, stems, roots) and the quality with which these are constructed. The approach taken by most GVMs for handling this functional diversity, is to consider only a few archetypal PFTs \citep{harrisonetal2010}, assumed to differ in key physiological attributes. While it is common to assume differences in biomass partitioning between PFTs, these estimates are based on limited empirical data. Moreover, there is little consensus on how biomass partitioning and allocation (see Methods \ref{sec:partterm}) should be modelled in GVMs \citep{franklin2012,dekauwe2014,friend2014}. These shortcomings largely reflect the lack of suitable datasets of global scope with which models can be tested, constrained and compared.

Here, we use a unique, new database\citep{Falster2015} (Fig. 1) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. The dataset includes over 21,000 observations of biomass and size metrics for individual plants, contributed by 175 studies across nine vegetation types (Fig. S1). Previous work based on whole stands\citep{oneilldeangelis1981,enquistniklas2002,reich2014} and very small plants\citep{poorter2012}, reported differences among angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). Here we extend this finding to show that LMF differs not only across the angiosperm-gymnosperm divide, but also within angiosperms, at the level of individual plants, and across the entire size range of woody plants in natural forests (0.1-100m) (Fig. 2a).

Moreover, we show that differences in LMF among woody PFTs arise due to differences in leaf morphology, not in the amount of leaf area held on the plant (Fig. 2b-c). Previously, it was unclear whether PFTs with higher leaf biomass also had higher total leaf area. Leaf area is relevant because it drives total light interception and thus photosynthesis. Several classic studies have assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}, suggesting a growth advantage. We found no evidence for this, instead we found that the three PFTS maintain similar leaf area per unit above-ground biomass (the `leaf area ratio`, LAR). Because of allometric relationships between the studied variables \citep{niklas1994}, there was a strong effect of plant height on LMF (Fig. 2a and S2). We corrected for plant height by estimating LMF at a common plant height, using a generalized additive model, a semi-parametric method that makes no assumption on the underlying functional relationships \citep{wood2006}. We then found that LMF at a common height was proportional to the leaf construction cost, as measured by leaf mass per area (LMA) (Fig. 2b). The consequence of this is that the leaf area ratio (the amount of leaf area supported per unit above-ground biomass), was invariant between PFTs, because $\textrm{LAR} = LMF / \textrm{LMA}$ (Fig. 2c and S3).

Another key metric governing vegetation behavior is the amount of leaf maintained per unit cross-sectional area of stem; quantifying the amount of light intercepting tissue relative to the amount of tissue available to transport water through the stem. This measure has long been assumed to ..(reason why it's important)... Mirroring the results for LMF, we find that the amount of leaf mass per unit stem area differs among the three PFTs, that these differences also relate to LMA (Fig. 2d), and that PFTs are similarly invariant in the amount of leaf area per unit stem area (Fig. 2e).

The above results highlight how differences in LMA -- a central trait of the leaf economics spectrum\citep{wrightetal2004} -- drives differences in biomass partitioning among dominant woody PFTs. Based on a XXX statistical model, we found that LMA alone could explain almost as much  (?more?) variation as PFT, even though LMA only captures aspects of leaf physiology. These results suggest promising avenues for parameterizing and simplifying GVMs.

Given these results, we were also interested to know whether additional variation in biomass partitioning could be explained by either climate, within PFT variation in LMA or by other traits. Despite the size of our dataset, the strong size-dependence of LMF greatly hinders comparisons among species and across climatic gradients, due to the small sample size within each species or site. Thus, to answer these additional questions we found it helpful to decompose LMF as the product, 
\begin{equation}\label{eq:M1}\textrm{LMF} = \left(\frac{M_F}{A_S}\right) \left(\frac{A_S}{M_T}\right).\end{equation}
An advantage of eq. \ref{eq:M1} is that the second term absorbs nearly all of the size related variation in LMF (Tables 1-2) and is otherwise fairly conserved between species\citep{chave2005} and across PFTs (Fig. S4). As shown in Fig. 2b and Fig. 2d, the remaining term $M_F/A_S$ exhibits comparable differences to LMF, but unlike $M_F/A_S$ is independent of plant height (Tables 1-2). $M_F/A_S$ is thus a useful of LMF that can be compared across species and climates.

We analyzed climate effects in two ways: with quantitative variables (mean annual precipitation, MAP; mean annual temperature, MAT), and by biome (tropical / temperate/ boreal), a simple classification taking into account both MAT and MAP (Fig. S1). Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig. 3). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table 2). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig. 3b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table 1).

The finding that climate does not appreciably affect biomass partitioning is good news for models that parameterize the globe by PFT. Yet, these results contrast those of a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for these differences is that our measurements were taken on individual plants, whereas in the other study they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig. S5), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT seem stronger than any direct climate effect on biomass partitioning within individual plants (Tables 1 and 2).

In contrast, we found that variation among species in leaf and stem traits could explain additional variation in biomass partitioning. The effect of LMA was found to hold also within PFTs (Fig. 4a), explaining 28\% of the variation in $M_F/A_S$ across species. The fact that PFTs differ in LMA is evident in the separation of the three groups along the x-axis of Fig. 4a. Yet these clouds were broad and overlapping, with a more general relationship extending across the entire LMA axis. The remaining (?? what is number??) variation in $M_F/A_S$  was explained by species-level differences in $A_F/A_S$ (Fig. 4b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig. 2e), but does vary close to two orders of magnitude across species. The separation among PFTs along the Y-axis arises due to differences in LMA.

Overall, our results establish general patterns about plant construction and thus lay an empirical base against which models can be benchmarked. We found that PFTs differed in LMF, and this was proportional to LMA. It thus seems possible to integrate a key leaf trait with whole-plant modelling of biomass partitioning. A recent study compared allocation routines in a number of leading ecosystem models \citep{dekauwe2014}, and recommended constraining allocation by observed biomass fractions instead of using constant allocation fractions. A first approximation within GVMs would be to keep leaf area to stem cross-sectional area constant for a given PFT, and vary biomass allocation accordingly. LMA, already a parameter in most GVMs, then gives the ratio of leaf biomass to stem biomass. An example of such a model is LPJ \citep{sitchetal2003}, where the ratio of leaf area to sapwood area is a constant parameter, but LMA differs between PFTs. More fundamentally, our results suggest PFTs defined on the basis of two key traits, LMA and $A_F/A_S$ would better capture patterns of biomass partitioning than existing groups, which only account for differences in one of the two traits identified here. In any case, the growing availability of large datasets on stand biomass and individual plant construction  \citep{Falster2015} suggest the time is ripe for rigorous benchmarking \citep{abramowitz_towards_2012, dekauwe2014} of GVMs against empirical data.

@@

We used the recently compiled Biomass And Allometry Database (BAAD) \citep{Falster2015}, including a total of \Sexpr{nrow(baad_all$data)} observations, but here we use only data that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_W$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (ca. 1.3m) ($A_{Sbh}$).  For each analysis, we used different subsets of the data. Sample sizes by PFT are summarized in Tables~\ref{table:count1}-\ref{table:count4} We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

$$\textrm{LMF} = \frac{M_F}{M_T} = \left(\frac{M_F}{A_F}\right) \left(\frac{A_F}{M_T}\right) = \textrm{LMA} \times \textrm{LAR}.$$

@@

$$\textrm{LMF} = \frac{M_F}{M_T} = \left(\frac{M_F}{A_S}\right) \left(\frac{M_T}{A_S}\right)^{-1}.$$

@@

R.A.D and D.S.F. conceived the ideas and methods, and co-wrote the paper. R.A.D. analysed data.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure3.pdf}
    \caption{\textbf{Beyond influencing the distribution of PFTs, climate only modestly influences biomass partitioning}. a) LMF for each PFT within three different biomes.  No leaf area estimates were available for boreal deciduous angiosperms. Error bars are 95\% confidence intervals for the mean, estimated from a linear mixed-effects model using species within study as the random effect. Letters denote significant differences (at $\alpha = 0.05$). b-c) Relationships between climate and LMF within each PFT. Solid lines are generalized additive model (GAM) fits (with a base dimension of 3); with grey areas indicating the 95\% confidence interval around the GAM.}
    \label{fig:figure3}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{Figure4.pdf}
     \caption{\textbf{Leaf and stem traits capture variation in biomass partitioning across species.} Individual plant data were averaged by species and study combinations. Lines are linear regressions with 95\% confidence intervals. Both regressions were significant ($P \textless 0.01$). $R^2$ for fitted relationships are 0.28 and XXX in the right panel.}
  \label{fig:figure4}
\end{figure}

@@

@@

@@

$$
\frac{dM_i}{dt} = a_i \textrm{NPP} - \tau_i M_i
$$
